Comparing characters

### 2016-12-20 Analysis/Character_difference/Character_diff_empirical.Rmd
Cleaned Analysis folder

```{r}
library(Claddis) ; library(dispRity)
source("functions.R") ; load.functions(test = FALSE)

## Load the C libraries
dyn.load("../Functions/char.diff.so")

@@

```{r}
## Set up the matrices path
matrix_path = "../Data/100_Matrices"
files <- list.files(path = matrix_path, pattern = ".nex")

>>>>

```{r}
library(Claddis) ; library(dispRity)
source("../../Functions/cor.matrix.R")
source("../../Functions/modify.matrix.R")
source("../../Functions/plot.char.diff.density.R")
source("../../Functions/plot.cor.matrix.R")
source("../../Functions/write.nexus.std.R")
source("../../Functions/char.diff.R")

## Load the C libraries
dyn.load("../../Functions/char.diff.so")

@@

```{r}
## Set up the matrices path
matrix_path = "../../Data/100_Matrices"
files <- list.files(path = matrix_path, pattern = ".nex")

### 2016-12-20 Analysis/Character_difference/Comparing_characters.Rmd
Cleaned Analysis folder

```{r}
cor.spearman <- function(X, Y) {
    return(1-round(abs(cor(X,Y, method = "spearman")), digit = 10))
}
## Note that to avoid rounding issues, we will round the correlation value up to the last 10 digits
```

We will also introduce a new difference simply called "Character difference" that counts the difference in states pattern (ranging from 0 - no difference - to 1 - complete difference) as follows:

@@

Note that in practice, the characters are first normalised and the Gower distance is then calculated.
The normalisation works by changing the characters tokens labels to always start from `0` being the first state.
For example, a character `{8,8,2,1,3,0}` will be normalised by changing all the occurences of the first state (`8`) by `0`, then all the occurences of the second one (`2`) by `1` and so on resulting in the character being normalised as `{0,0,1,2,3,4}`.
Note that from a phylogenetic point of view, if the character is Fitch-like, both characters convey the exact same information.

```{r}
source("../Functions/char.diff.R")
dyn.load("../Functions/char.diff.so")
```

A simple case
----------------
Let's create a bunch of binary characters first

@@

Straight away, we expect `A` and `B` to be totally correlated as well as `D` and `E`.
Also, will will check if our difference metric is triangular (i.e. if `A` = `B` then the difference `(A,X)` must be equal to `(B,X)`).

@@

A non-binary case
----------------
However, in reality, some characters might have more than two state.

@@

In this case again, we expect `A` and

Straight away, we expect `A` and `B` to be totally correlated/not different as well as well as `C` and `D`.
However, the spearman correlation does pick up some difference (i.e. not totally correlated)

@@

```{r}
source("../Functions/cor.matrix.R")
pairwise_matrix_correlation <- cor.matrix(matrix, cor.spearman)
pairwise_matrix_differences <- char.diff(matrix)
```

@@

```{r, fig.width=6, fig.height=6}
source("../Functions/plot.cor.matrix.R")
## Extracting the first character differences
character_differences <- pairwise_matrix_differences[,1]
character_correlation <- pairwise_matrix_correlation[,1]

>>>>

```{r}
cor.spearman <- function(X, Y) {
    return(1-round(abs(cor(X,Y, method = "spearman")), digit = 10))
}
```
> Note that to avoid rounding issues, we will round the correlation value up to the last 10 digits

We will also introduce a new difference called "Character Difference" that counts the difference in states pattern (ranging from 0 - no difference - to 1 - complete difference) as follows:

@@

Note that in practice, the characters are first normalised and the Gower distance is then calculated:
 * The normalisation works by changing the characters tokens labels to always start from `0` being the first state.
For example, a character `{8,8,2,1,3,0}` will be normalised by changing all the occurrences of the first state (`8`) by `0`, then all the occurrences of the second one (`2`) by `1` and so on resulting in the character being normalised as `{0,0,1,2,3,4}`.
Note that from a phylogenetic point of view, if the character is Fitch-like, both characters convey the exact same information.
 * The Gower distance is simply the distance divided by the number of observation (i.e. to take into account the inapplicable or missing data).

The function to calculate the character difference is a modification of the `stats::dist` function's source code (in `C`):

```{r}
source("../../Functions/char.diff.R")
dyn.load("../../Functions/char.diff.so")
```

A simple case
----------------
Let's create a bunch of binary characters first:

@@

Straight away, we expect `A` and `B` to be totally correlated as well as `D` and `E`.
Also, we will check if our difference metric is triangular (i.e. if `A` = `B` then the difference `(A,X)` must be equal to `(B,X)`).

@@

A non-binary case
----------------
However, in reality, some characters might have more than two state:

@@

Similarly, we expect `A` and `B` to be totally totally correlated as well as `C` and `D`.
However, the Spearman correlation does pick up some difference (i.e. not totally correlated)

@@

```{r}
source("../../Functions/cor.matrix.R")
pairwise_matrix_correlation <- cor.matrix(matrix, cor.spearman)
pairwise_matrix_differences <- char.diff(matrix)
```

@@

```{r, fig.width=6, fig.height=6}
source("../../Functions/plot.cor.matrix.R")
## Extracting the first character differences
character_differences <- pairwise_matrix_differences[,1]
character_correlation <- pairwise_matrix_correlation[,1]
